Storm Data Analysis - Week 4

### 2021-03-22 Manuscript.Rmd
- Selected 95th percentile of overall data - Tests for panel plot

@@

# Make a copy of data and cast as DT
dt_data <- as.data.table(lean_data)
# Define the number of rows that will be in the grouped DT. This is equal to the
# unique values of event, as we'll group by that.
denominator <- uniqueN(dt_data, "event")
# Apply final transformations to DT
sel_data <- dt_data[, .(
                        counter = .N, # Count elements in group
                        # Total monetary damages per group
                        total_dmg = sum(tot_dmg, na.rm = TRUE),
                        # Total fatalities
                        fatalities = sum(FATALITIES, na.rm = TRUE),
                        # Injuries
                        injuries = sum(INJURIES, na.rm = TRUE)
                        ),
                    # Grouping by event type
                    by = event][
                        ,
                        # Compute a "percent rank": rank events by number of
                        # occurrences (desc), then assign a percentile by
                        # dividing rank by denominator (i.e. nrows)
                        # Note double dot syntax to access calling scope
                        `:=`(perc_rank = frankv(counter)/..denominator,
                             # Same for monetary damage
                             dmg_rank = frankv(total_dmg)/..denominator,
                             # And life damage
                             inj_rank = frankv(injuries)/..denominator,
                             fat_rank = frankv(fatalities)/..denominator
                             )
                        ][
                            # Filter for frequency rank >=90th
                            perc_rank >= 0.9
                        ][
                            order(-counter)
                        ]

@@

>>>>

# Min-max normalization
normalize <- function(x) {
    return ((x - min(x)) / (max(x) - min(x)))
  }

@@

# Make a copy of data and cast as DT
dt_data <- as.data.table(lean_data)
# Define the number of rows that will be in the grouped DT. This is equal to the
# unique values of event, as we'll group by that.
denominator <- uniqueN(dt_data, "event")
# Apply final transformations to DT
sel_data <- dt_data[, .(
                        counter = .N, # Count elements in group
                        # Total monetary damages per group
                        total_dmg = sum(tot_dmg, na.rm = TRUE),
                        # Total fatalities
                        fatalities = sum(FATALITIES, na.rm = TRUE),
                        # Injuries
                        injuries = sum(INJURIES, na.rm = TRUE)
                        ),
                    # Grouping by event type
                    by = event][
                        ,
                        # Compute a "percent rank": rank events by number of
                        # occurrences (desc), then assign a percentile by
                        # dividing rank by denominator (i.e. nrows)
                        # Note double dot syntax to access calling scope
                        `:=`(perc_rank = frankv(counter)/..denominator,
                             # Same for monetary damage
                             dmg_rank = frankv(total_dmg)/..denominator,
                             # And life damage
                             inj_rank = frankv(injuries)/..denominator,
                             fat_rank = frankv(fatalities)/..denominator
                             )
                        ][
                            # Filter for frequency rank >=95th
                            perc_rank >= 0.95
                        ][
                            order(-counter)
                        ]

@@

Figure 1 is a panel plot of data at or above the 95^th^ percentile of (A) event frequency by type, (B) monetary damages and (C) damages to persons in the period of interest (1950–2013). One can easily that the magnitude of event-related damages decreases almost exponentially with frequency. It should be noted that the preponderant majority of events below 95^th^ percentile of frequency are actually data-entry errors arising from "event type" being an open-text field in the database.

```{r frequency-plots}

sel_data %>% ggplot(
    aes(x = fct_reorder(event, injuries, .desc = T), y = inj_rank)
    ) +
    geom_point()
